---
title: "gp"
author: "Ty Wicks"
date: "4/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Project #2 USGS DOC Data


######################
#Install Libraries
######################


```{r}
library(fields)
library(lattice)
library(glmnet)
```


######################
#Load data
######################


```{r}
load("DOC_baseFlow_weightedAvg_and_predictors.RData")

#This is a really long name for a dataset, so this command just renames it.
data <- DOC_baseFlow_weightedAvg_and_predictors
dim(data)
colnames(data)
lm(meanDOC ~., data = data)

y_response <- data[,"meanDOC"]

year.obs <- table(data$Year)
plot(1984:2012, year.obs, type="b", pch=19, xlab="Year", ylab="Number of Observations")
barchart(year.obs, horizontal=FALSE, col="gray", xlab="Year", ylab="Number of Observations")
```


######################
#Plot the locations
######################


```{r}
plot(data$Longitude, data$Latitude, pch=19, col=rgb(0,0,0,.1), xlim=c(-109,-102), ylim=c(37,42))
US(add=T)
text(-108, 37.5, "Colorado", cex=2)
```


######################
#Creating Matrix of Predictors
######################


```{r}
#Aspect Ratio: indicates the predominant aspect, which is a categorical value with nine different classes: North, Northeast, East, Southeast, South, Southwest,West, Northwest, and Flat

xFactors <- model.matrix(y_response ~ as.factor(data[,"aspect_predominant"]))[, -1]

colNames_xFactors <- c("aspect_predominant_1")
for (index in 2:8)
{
  colNames_xFactors <- c(colNames_xFactors,paste("aspect_predominant_",as.character(index),sep=""))
}

colnames(xFactors) <- colNames_xFactors

#Collecting the aspect ratio variables with all of the others needed
#Using one fire predictor  (column numbers for fire predictors are 6, 7, and 8)
X_predictors_unStandardized <- as.matrix( cbind( DOC_baseFlow_weightedAvg_and_predictors[,c(2,6,9:30,32)], xFactors) )

head(X_predictors_unStandardized)
dim(X_predictors_unStandardized)
class(X_predictors_unStandardized)
dataFrame_X_predictors <- as.data.frame(X_predictors_unStandardized)
```


######################
#Standardizing the Predictors 
#(Generally a good idea before fitting a linear model)
######################


```{r}
X_predictors_means <- colMeans(X_predictors_unStandardized)
X_predictors_sd <- apply(X_predictors_unStandardized, MARGIN=2, 'sd')

time_index <- 			1
fire_indices <- 		2
MPB_indecies <- 		3:5
landCover_indices <- 	6:8
temp_indices <- 		9:14
precipSnow_indices <- 	15:20
soil_indices <- 		21:23
elevation_index <- 		24
wasteWater_index <- 	25
aspec_indices <- 		26:33
X_indices_subtractMean <- c(time_index, temp_indices, precipSnow_indices, soil_indices, elevation_index, wasteWater_index)
X_indices_scaleBySD <- c(fire_indices)

X_predictors_standardized <- X_predictors_unStandardized
X_predictors_standardized[,X_indices_subtractMean] <- sweep(X_predictors_unStandardized[,X_indices_subtractMean],MARGIN=2,X_predictors_means[X_indices_subtractMean],'-')
  
colnames(X_predictors_standardized)
dataFrame_X_standardized <- as.data.frame(X_predictors_standardized)
View(dataFrame_X_standardized)
```


######################
#Create new variables
#Plot all data points with DOC
######################


```{r}
doc_data <- cbind(y_response, dataFrame_X_standardized)

doc_fit <- lm(y_response ~ ., data = doc_data)
summary(doc_fit)

col_names <- colnames(doc_data)

for(var in 2:ncol(doc_data)){
  plot(doc_data[ , var], doc_data$y_response, ylab = "DOC", xlab = col_names[var])
  title(paste0(col_names[var], " by DOC"))
  
  abline(lm(y_response ~ doc_data[ , var], data = doc_data), col = "red")
}
```


######################
#Find best variables using
#Forward & Backward step and LASSO
######################


```{r}
stepBck <- step(doc_fit, direction="backward", k = log(nrow(doc_data)), trace = 0)

fwdFit <- lm(y_response ~ 1, data = doc_data)
stepFwd <- step(fwdFit, direction="forward", k = log(nrow(doc_data)), scope = formula(doc_fit), trace = 0)

remove <- (-which(colnames(doc_data) == 'y_response'))
lassMat <- as.matrix(doc_data[,remove])
cvLass <- cv.glmnet(lassMat, doc_data$y_response)
```


######################
#Review method summaries and
#divide results into groups
######################


```{r}
summary(stepBck)
#fire_normRdNBR_5yDecay, landCover_developedFraction, precip_baseFlow, SWE_baseFlow, soil_Bfraction, elevation_mean,
#wasteWaterPointSources_count, aspect_predominant_3

summary(stepFwd)
#elevation_mean, wasteWaterPointSources_count, aspect_predominant_3, landCover_developedFraction, soil_Bfraction,
#fire_normRdNBR_5yDecay, aspect_predominant_7

coef(cvLass)
#maxTemp_yearAvg, elevation_mean, wasteWaterPointSources_count, aspect_predominant_3


#Chosen 3 times:
#elevation_mean, wasteWaterPointSources_count, aspect_predominant_3

#Chosen 2 times:
#landCover_developedFraction, soil_Bfraction, fire_normRdNBR_5yDecay

#Chosen 1 time:
#precip_baseFlow, SWE_baseFlow, aspect_predominant_7, maxTemp_yearAvg
```


######################
#Create new data frame of
#best variables for graphing
######################


```{r}
chosenData <- doc_data[,c(25, 26, 29, 33, 7, 23, 3, 16, 15, 19)]  

chosenVars <- colnames(chosenData)
```


######################
#Graph each variable
#side by side
######################


```{r}
par(mfrow=c(1,5))
plot(chosenData$elevation_mean, doc_data$y_response, xlab = "Elevation", ylab = "DOC", main = "3 Time Selection")
plot(chosenData$wasteWaterPointSources_count, doc_data$y_response, xlab = "Waste Water Source", ylab = "DOC", main = "3 Time Selection")
#plot(chosenData$aspect_predominant_3, doc_data$y_response, xlab = "Aspect 3", ylab = "DOC", main = "3 Time Selection")
plot(chosenData$landCover_developedFraction, doc_data$y_response, xlab = "Land Development Fraction", ylab = "DOC", main = "2 Time Selection")
plot(chosenData$soil_Bfraction, doc_data$y_response, xlab = "Soil BFraction", ylab = "DOC", main = "2 Time Selection")

par(mfrow=c(1,5))
plot(chosenData$fire_normRdNBR_5yDecay, doc_data$y_response, xlab = "Fire Norm Decay", ylab = "DOC", main = "2 Time Selection")
plot(chosenData$precip_baseFlow, doc_data$y_response, xlab = "Precipitation Base Flow", ylab = "DOC", main = "1 Time Selection")
plot(chosenData$SWE_baseFlow, doc_data$y_response, xlab = "SWE Base Flow", ylab = "DOC", main = "1 Time Selection")
#plot(chosenData$aspect_predominant_7, doc_data$y_response, xlab = "Aspect 7", ylab = "DOC", main = "1 Time Selection")
plot(chosenData$maxTemp_yearAvg, doc_data$y_response, xlab = "Yearly Max Temp", ylab = "DOC", main = "1 Time Selection")
```


######################
#Graph variables chosen
#by all three methods
######################


```{r}
par(mfrow=c(1,2))
plot(chosenData$elevation_mean, doc_data$y_response, xlab = "Elevation", ylab = "DOC", main = "3 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$elevation_mean), col = "red")
plot(chosenData$wasteWaterPointSources_count, doc_data$y_response, xlab = "Waste Water Source", ylab = "DOC", main = "3 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$wasteWaterPointSources_count), col = "red")
```


######################
#Graph variables chosen
#by two methods
######################


```{r}
par(mfrow=c(1,3))
plot(chosenData$landCover_developedFraction, doc_data$y_response, xlab = "Land Development Fraction", ylab = "DOC", main = "2 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$landCover_developedFraction), col = "red")
plot(chosenData$soil_Bfraction, doc_data$y_response, xlab = "Soil BFraction", ylab = "DOC", main = "2 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$soil_Bfraction), col = "red")
plot(chosenData$fire_normRdNBR_5yDecay, doc_data$y_response, xlab = "Fire Norm Decay", ylab = "DOC", main = "2 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$fire_normRdNBR_5yDecay), col = "red")
```


######################
#Graph variables chosen
#by one method
######################


```{r}
par(mfrow=c(1,3))
plot(chosenData$precip_baseFlow, doc_data$y_response, xlab = "Precipitation Base Flow", ylab = "DOC", main = "1 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$precip_baseFlow), col = "red")
plot(chosenData$SWE_baseFlow, doc_data$y_response, xlab = "SWE Base Flow", ylab = "DOC", main = "1 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$SWE_baseFlow), col = "red")
plot(chosenData$maxTemp_yearAvg, doc_data$y_response, xlab = "Yearly Max Temp", ylab = "DOC", main = "1 Time Selection")
abline(lm(doc_data$y_response ~ chosenData$maxTemp_yearAvg), col = "red")
```




